---
title: "Scenarios_for_GFI"
author: "Natalie Rice"
date: "12/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Base input 
input_1 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#Answer was go for it 
```

```{r}
#Changing yardline_100 to 50 
input_2 <- data.frame(
  "yardline_100" = 50,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#Answer was go for it 
```

```{r}
#Yardline_100 to 80

input_3 <- data.frame(
  "yardline_100" = 80,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#Answer was don't go for it
```

```{r}
#Total home score to 14 (now teams are tied with 14)

input_4 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#Answer was go for it 
```


```{r}
#Total home score to 14 (now teams are tied with 14) but further from endzone

input_5 <- data.frame(
  "yardline_100" = 50,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#go for it 
```

```{r}
#Total away score to 27 (away team beating home team by 7)

input_6 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#Don't go for it 

```

```{r}
#Team with possession has 1 timeout, defending team has 2

input_7 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 1,
  "defteam_timeouts_remaining" = 2
)

#go for it 
```

```{r}
#Team with possession has 3 timeouts, defending team has 2

input_8 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 3,
  "defteam_timeouts_remaining" = 2
)

#go for it 
```

```{r}
#Team with possession has 0 timeouts, defending team has 2

input_9 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 0,
  "defteam_timeouts_remaining" = 2
)

#go for it 
```

```{r}
#Team with possession has 2 timeouts, defending team has 1

input_10 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 1
)

#go for it 
```

```{r}
#Team with possession has 2 timeouts, defending team has 3

input_11 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 3
)

#go for it 
```

```{r}
#Team with possession has 2 timeouts, defending team has 0

input_12 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 0
)

#go for it 
```

```{r}
#Yards to go changed to 2

input_13 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 2,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#go for it 
```

```{r}
#Yards to go changed to 7
input_14 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = 'Half1',
  "goal_to_go" = 0,
  "ydstogo" = 7,
  "play_type" = 'run',
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)

#go for it 
```

```{r}
#half second remaining 600 in first half

input_15 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#don't go for it 
```

```{r}
#seconds remaining 600 in second half

input_16 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#go for it 
```

```{r}
#half second remaining 120 in the first half

input_17 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#go for it 
```

```{r}
#Base input but with timeout = 1 (called a timeout)

input_18 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 1,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#go for it 
```

#Changing time remaining and home team losing 
```{r}
#second half remaining 600 in second half and home team losing
input_19 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#go for it 
```

```{r}
#remaining 600 seconds in first half and home team losing 
input_20 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)
#go for it 
```

```{r}
#half second remaining 120 in the first half and home team losing 
input_21 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#go for it
```

#Changing time remaining and teams are tied 
```{r}
#remaining 600 seconds in first half and tied
input_22 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#go for it 
```

```{r}
#600 seconds remaining in second half
input_23 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)
#go for it 
```

```{r}
#2 min remaining 120 in the first half
input_24 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#go for it 
```





#Play begins with posession team in shotgun position
```{r}
#Base input with shotgun
input_25 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = 'Half1',
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = 'run',
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 21, 
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 600 in first half and home team losing, and shotgun
input_26 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 600 in second half and home team losing, and shotgun
input_27 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 120 in the first half and home team losing, and shotgun 
input_28 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

#Changing time remaining and teams are tied 
```{r}
#half second remaining 600 in first half, and shotgun
input_29 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 600 in second half, and shotgun
input_30 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 120 in the first half, and shotgun
input_31 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

#Repeat all above simulations with New England Patriots (NE)

```{r}
#Base input 
input_32 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21, 
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#Changing yardline_100 to 50 

input_33 <- data.frame(
  "yardline_100" = 50,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#Yardline_100 to 80

input_33 <- data.frame(
  "yardline_100" = 80,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#Total home score to 14 (now teams are tied with 14)

input_34 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```


```{r}
#Total home score to 14 (now teams are tied with 14) but further from endzone

input_35 <- data.frame(
  "yardline_100" = 50,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#Total away score to 27 (away team beating home team by 7)

input_36 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)

```

```{r}
#Team with possession has 1 timeout, defeding team has 2
input_37 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 1,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#Team with possession has 3 timeouts, defending team has 2
input_38 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 3,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#Team with possession has 0 timeouts, defending team has 2
input_39 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 0,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#Team with possession has 2 timeouts, defending team has 1
input_40 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 1 
)
```

```{r}
#Team with possession has 2 timeouts, defending team has 3
input_41 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 3 
)
```

```{r}
#Team with possession has 2 timeouts, defending team has 0

input_42 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 0 
)
```

```{r}
#Yards to go changed to 2
input_43 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 2,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#Yards to go changed to 7
input_44 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 7,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 600 in first half
input_45 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 600 in second half
input_46 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 120 in the first half
input_47 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#Base input but with timeout = 1 (called a timeout)
input_48 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 1,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

#Changing time remaining and home team losing 
```{r}
#half second remaining 600 in first half and home team losing
input_49 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 600 in second half and home team losing 
input_50 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 120 in the first half and home team losing 
input_51 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

#Changing time remaining and teams are tied 
```{r}
#half second remaining 600 in first half
input_52 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 600 in second half
input_53 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 120 in the first half
input_54 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

#Play begins with posession team in shotgun position
```{r}
#Base input with shotgun
input_55 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 21, 
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 600 in first half and home team losing, and shotgun
input_56 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 600 in second half and home team losing, and shotgun
input_57 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 120 in the first half and home team losing, and shotgun 
input_58 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 27,
  "away_score" = 27,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

#Changing time remaining and teams are tied 
```{r}
#half second remaining 600 in first half, and shotgun
input_59 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 600 in second half, and shotgun
input_60 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 600,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = "run",
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}
#half second remaining 120 in the first half, and shotgun
input_61 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 5,
  "play_type" = 'run',
  "shotgun" = 1,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 14,
  "away_score" = 14,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "NE",
  "posteam" = "NE",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2 
)
```

```{r}

#original but changed yds to go to  1 
input_62 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 1,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 21,
  "defteam_score" = 3,
  "away_score" = 3,
  "home_score" = 21,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 3,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#answer was go for it 
```

```{r}

#Lions vs Vikings Scenario
input_63 <- data.frame(
  "yardline_100" = 72,
  "half_seconds_remaining" = 248,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 1,
  "play_type" = "pass",
  "shotgun" = 0,
  "timeout" = 1,
  "posteam_score" = 23,
  "defteam_score" = 21,
  "away_score" = 21,
  "home_score" = 23,
  "season" = 2021,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 4,
  "posteam_timeouts_remaining" = 1,
  "defteam_timeouts_remaining" = 3
)

#answer was go for it 
```


```{r}

#Lions vs Vikings Scenario
input_64 <- data.frame(
  "yardline_100" = 99,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 2,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 21 ,
  "away_score" = 21,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 4,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#answer was go for it 
```


```{r}

#Lions vs Vikings Scenario
input_65 <- data.frame(
  "yardline_100" = 63,
  "half_seconds_remaining" = 120,
  "game_half" = "Half1",
  "goal_to_go" = 0,
  "ydstogo" = 2,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 21 ,
  "away_score" = 21,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 4,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#answer was go for it 
```

```{r}

#Lions vs Vikings Scenario
input_66 <- data.frame(
  "yardline_100" = 10,
  "half_seconds_remaining" = 120,
  "game_half" = "Half2",
  "goal_to_go" = 0,
  "ydstogo" = 3,
  "play_type" = "run",
  "shotgun" = 0,
  "timeout" = 0,
  "posteam_score" = 14,
  "defteam_score" = 21 ,
  "away_score" = 21,
  "home_score" = 14,
  "season" = 2018,
  "home_team" = "DET",
  "posteam" = "DET",
  "roof" = "dome",
  "down" = 4,
  "posteam_timeouts_remaining" = 2,
  "defteam_timeouts_remaining" = 2
)

#answer was go for it 
```
